---
title: "Template"
author: "`r system('whoami', intern = TRUE)`"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
    toc: true
    toc_float: true
    collapsed: true
    smooth_scroll: true
    df_print: paged
    code_download: true
    highlight: kate
    code_folding: hide
editor_options:
  chunk_output_type: console
params:
  fig_path: "figures/"
  d_out: "./"
---

# Preparations

## Set global options

```{r style, echo = FALSE, results="asis", cache=FALSE, message=FALSE}
# v.05

# This piece of code is necessary if running Rmd on vscode,
# which does not read the header
if (!exists("params")) {
  params <- list(fig_path = "figures/", d_out = "./")
}

# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(
  # General chunk configuration
  cache = FALSE,
  warning = FALSE, # Make warning FALSE, unless explicitly turn it on
  message = TRUE, # Turn to FALSE to avoid verbose and to have clean documents
  echo = TRUE,
  tidy = FALSE,
  include = TRUE,
  # Figure configurations
  fig.path = params$fig_path,
  dev = c("png"),
  fig.height = 5,
  fig.width = 5 * golden_ratio,
  fig.wide = TRUE,
  out.width = "150%", # Make figures appear larger and fit to the page width
  dpi = 300,
  comment = "  "
)

# Set seed for reproducibility
set.seed(13)

# Set plot options
library("ggplot2")
library(ggthemes)
theme_set(theme_bw(base_size = 10))

#  Set color blind friendly pallet as default for discrete colors
# Source of palette: https://www.sthda.com/english/articles/32-r-graphics-essentials/125-ggplot-cheat-sheet-for-great-customization #nolint
cbp1 <- c(
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442",
  "#0072B2",
  "#D55E00",
  "#CC79A7",
  "#999999"
)

dynamic_cbp1_colors <- function(n) {
  colorRampPalette(cbp1)(n)
}

# Define custom scales using the dynamic color generator
scale_fill_custom <- function(...) {
  discrete_scale(
    aesthetics = "fill",
    scale_name = "custom",
    palette = dynamic_cbp1_colors,
    ...
  )
}
scale_colour_custom <- function(...) {
  discrete_scale(
    aesthetics = "colour",
    scale_name = "custom",
    palette = dynamic_cbp1_colors, ...
  )
}

# Set the custom scales as default
options(
  ggplot2.discrete.fill = scale_fill_custom,
  ggplot2.discrete.colour = scale_colour_custom
)

# Set viridis as default for scale continuous
options(
  ggplot2.continuous.fill = scale_fill_viridis_c,
  ggplot2.continuous.colour = scale_colour_viridis_c
)


#' Display Data Frame with Formatted Numeric Columns
#'
#' This function formats a data frame by rounding numeric columns and
#' displaying it either as a static table using `knitr::kable` or as an
#' interactive table using `DT::datatable`.
#'
#' @param data A data frame to be formatted and displayed.
#' @param type A character string specifying the display type: "static"
#' for a static table using `knitr::kable`,
#' "interactive" for an interactive table using `DT::datatable`.
#' Default is "static".
#' @param n_round An integer specifying the number of decimal places to round
#'  numeric columns. Default is 3.
#'
#' @return Formatted data frame displayed as specified.
#' @examples
#' \dontrun{
#' show_df(data, type = "static")
#' show_df(data, type = "interactive", n_round = 2)
#' }
show_df <- function(data, type = "static", n_round = 3) {
  data <- data %>% mutate_if(is.numeric, \(x) round(x, n_round))
  if (type == "static") {
    data <- knitr::kable(data, row.names = FALSE)
  }
  if (type == "interactive") {
    data <- DT::datatable(data,
      rownames = FALSE, extensions = "FixedColumns",
      options = list(scrollX = TRUE, fixedColumns = list(leftColumns = 1))
    )
  }
  data
}
# Set output directory
d_out <- params$d_out
rm(params)
```


## Libraries

```{r}
library(tidyverse)
```
# Introduction

I will demonstrate how to use this template.

I normally knit this script using the axiliary script `knit_conda` (for R on conda) or `knit_local` (for R installed locally). If it does not work, check if paths (pandoc) and assumptions in the script are met. 

```{bash, eval = FALSE}
bash knit_local 00_template.Rmd
```
If your code is in nested somewhere in the `code` folder, the output will be in the corresponding folder in `results`. If not, the output will be in the `working directory`.


>Of note, the figures are saved in the `figures` folder as `png`.

# Input data
Input data is just a chatGPT request on Tarantino's movies. This data is as fake as a two-dollar bill.


```{r}
# Create the data frame using tribble
tarantino_scenes <- tribble(
  ~movie, ~salary_per_movie, ~actor, ~minutes_on_screen,
  "Pulp Fiction", 56000, "Samuel L. Jackson", 45,
  "Kill Bill: Vol. 1", 60000, "Uma Thurman", 60,
  "Kill Bill: Vol. 2", 55000, "Christoph Waltz", 30,
  "Inglourious Basterds", 70000, "Brad Pitt", 50,
  "Django Unchained", 80000, "Leonardo DiCaprio", 40,
  "The Hateful Eight", 50000, "Kurt Russell", 55,
  "Once Upon a Time in Hollywood", 58000, "Jennifer Lawrence", 35,
  "Reservoir Dogs", 45000, "Harvey Keitel", 25,
  "Jackie Brown", 42000, "Pam Grier", 20,
  "Death Proof", 48000, "Michael Madsen", 15
)

# Print the table
glimpse(tarantino_scenes)
```

# Inspect data

## Static table
Use function `show_df` to display the data frame statically. This is based on `knitr::kable`.

```{r}
tarantino_scenes %>% show_df("static")
```

## Interactive table
Use function `show_df` to display the data frame interactively. This is based on `DT::datatable`.

```{r}
tarantino_scenes %>% show_df("interactive")
```

# Plot data {.tabset}

I modified the color of the plots. Note that the subsections here are in a tab.

## Discrete colors

Discrete colors are now based on a color blind friendly palette. You can modify it, by changing the `cbp1` vector in the header of the markdown.
```{r}
tarantino_scenes %>%
  ggplot(aes(
    x = fct_reorder(actor, desc(minutes_on_screen)),
    y = minutes_on_screen
  )) +
  geom_bar(stat = "identity", position = "dodge", aes(fill = actor)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Minutes on screen by actor",
    x = "Movie",
    y = "Minutes on screen"
  )
```

## Discrete colors with ordinal data

Ordinal data has ggplot original colors, which are viridis.
```{r}
tarantino_scenes %>%
  # Make actors ordinal
  mutate(actor = factor(actor, levels = unique(actor), order = TRUE)) %>%
  ggplot(aes(
    x = fct_reorder(actor, desc(minutes_on_screen)),
    y = minutes_on_screen
  )) +
  geom_bar(stat = "identity", position = "dodge", aes(fill = actor)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Minutes on screen by actor",
    x = "Movie",
    y = "Minutes on screen"
  )
```

## Continuous

Continuous colors are based on the viridis palette.

```{r}
tarantino_scenes %>%
  ggplot(aes(
    x = minutes_on_screen,
    y = salary_per_movie
  )) +
  geom_point(aes(color = salary_per_movie), size = 10)
```


# Conclusion

This is a template for Rmarkdown. I think it is pretty and useful.

# Perspectives

I would like to have it in Quarto, as it seems to become standard. However, I am having some issues for knitting in other folder with Quarto. So, for now, I will stick with Rmarkdown.

# Session info

```{r}
sessionInfo()
```
